Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-02-15 numerical-timestep.tex
working on timesteps.

@@

@@

\begin{equation}
\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L 
\label{eqn:dtaccel}
\end{equation}

 In equations~\ref{eqn:dthydro}-\ref{eqn:dtaccel}, the $min ( \ldots )_L$
formalism means that this value is calculated for all cells on a given level
L and the minimum overall value is taken as the timestep.  
Equation~\ref{eqn:dthydro} ensures that all cells satisfy the 
Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of an explicit
finite difference discretization of the Euler equations.  Effectively this condition
forces the timestep to be small enough such that any changes in the fluid propagate
less than a single grid spacing, $\Delta x$.  In this equation, $\kappa_{hydro}$ is 
a numerical constant with a value of $0 < \kappa_{hydro} \leq 1$ (with a typical
value of $\kappa_{hydro} \sim 0.3-0.5$) that ensures that the CFL condition is always
met, and $c_s$ and $v_x$ are the sound speed and peculiar baryon
velocity in a given cell.

Equation~\ref{eqn:dtaccel} is supplementary to equation~\ref{eqn:dthydro} in that it
takes into account the possibility of large accelerations causing numerical 
instabilities by violating the Courant condition.  In this equation, $\vec{g}$ is the
gravitational acceleration in each cell on level L.  \blue{ \emph{It should be
  noted that the timestep is computed before the acceleration field is computed,
  and the acceleration field is deleted at the end of each time step.
  So this condition isn't ever actually enforced.  We should either
  fix this problem, or not discuss it.
 -dcollins.} }

@@

>>>>

\begin{equation}
\Delta t_{MHD} = min \left( \kappa_{MHD} \frac{a \Delta x}{v_{f} + |v_x|} \right)_L ,
\label{eqn:dtMHD}
\end{equation}

@@

\begin{equation}
\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L ,
\label{eqn:dtaccel}
\end{equation}

\begin{equation}
\Delta t_{rad} = min \left(  \sqrt{\frac{\Delta x}{\vec{a}_{rad}} \right)_L,
\label{eqn:dtrad}
\end{equation}

\begin{equation}
\Delta t_{cond} = min \left(  \kappa_{cond} \frac{\Delta x^2
    n}{\kappa(T)} \right)_L,
\label{eqn:dtcond}
\end{equation}

@@

% \begin{equation}
% \Delta t_{} = 
% \label{eqn:dt}
% \end{equation}

% \begin{equation}
% \Delta t_{} = 
% \label{eqn:dt}
% \end{equation}

 In equations~\ref{eqn:dthydro}-\ref{eqn:dtcond}, the $min ( \ldots
)_L$ formalism means that this value is calculated for all cells on a
given level L and the minimum overall value is taken as the timestep.
Equation~\ref{eqn:dthydro} ensures that all cells satisfy the
Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of
an explicit finite difference discretization of the Euler equations.
Effectively this condition forces the timestep to be small enough such
that any changes in the fluid propagate less than a single grid
spacing, $\Delta x$.  In this equation, $\kappa_{hydro}$ is a
numerical constant with a value of $0 < \kappa_{hydro} \leq 1$ (with a
typical value of $\kappa_{hydro} \sim 0.3-0.5$) that ensures that the
CFL condition is always met, and $c_s$ and $v_x$ are the sound speed
and peculiar baryon velocity in a given cell.

@@

Equation~\ref{eqn:dtMHD} is only enforced when the equations of
magnetohydrodynamics are being solved, and is directly analogous to
Equation~\ref{eqn:dthydro} in that it ensures that the CLF condition
is being enforced at all times.  In this equation, $\kappa_{MHD}$ is a
numerical constant with a value of $0 < \kappa_{MHD} \leq 1$ (with a
typical value of $\kappa_{MHD} \sim 0.5$) that ensures that the CFL
condition is always met, and $v_f$ and $v_x$ are the ``fast wave
speed'' and peculiar baryon velocity in a given cell.  The fast wave
speed comes from a stability analysis of the MHD equations, and is
expressed as:

\begin{equation}
v_f = \sqrt{ \frac{1}{2} \left(  v_A^2 + c_s^2 + \sqrt{(v_A^2 +
      c_s^2)^2 - 4 v_A^2 c_s^2}  \right)  },
\label{eqn:vfastmhd}
\end{equation}

Where $c_s$ is the sound speed and $v_A$ is the Alfven speed, calculated
as $v_A = \sqrt{B^2/\rho_B}$ in units where $\mu_0 = 1$.

Equation~\ref{eqn:dtaccel} and~\ref{eqn:dtrad} are supplementary to equation~\ref{eqn:dthydro} in that they
take into account the possibility of large accelerations due to either
gravity (Eqn.~\ref{eqn:dtaccel} or radiation pressure
(Eqn.~\ref{eqn:dtrad}), which may cause numerical 
instabilities by violating the Courant condition.  In Equation~\ref{eqn:dtaccel}, $\vec{g}$ is the
gravitational acceleration in each cell on level L.  In
Equation~\ref{eqn:dtrad}, $\vec{a}_{rad}$ is the estimated
acceleration due to radiation pressure in each cell on level L,
defined as

\begin{equation}
\vec{a}_{rad} = \frac{ \sum_i \frac{\dot{E_i}}{c} \hat{r_i} }{m_b} 
\end{equation}

Where the sum calculates the energy deposited in a cell during the
previous timestep due to \textit{all photon packets} that crossed that
cell, with $\hat{r_i}$ being a unit vector that accounting for the
packet direction.

Equation~\ref{}
